### 2018-03-23 paper/sexSelectionSpacing-ver1.tex
References on spacing and maternal health

Parents' spacing between births has long served as a measure of son preference 
\citep{Leung1988}.
Before prenatal sex determination became available, the only recourse for 
parents who wanted a son---but did not yet have one---was to have the next 
birth sooner.
Son preference is therefore often associated with shorter spacing after the 
births of girls than boys 
\citep{Das1987,Rahman1993,Pong1994,Haughton1996,Arnold1997,Soest2012,Rossi2015}.
Shorter spacing is, in turn, associated with worse health outcomes for girls 
\citep{arnold98,Whitworth2002,Rutstein2005,Conde-Agudelo2006}.%
\footnote{
Parents are also more likely to cease childbearing after the birth of 
a son than after a daughter 
\citep{ben-porath76b,Das1987,Arnold1997,clark00,filmer09,Basu2010,Altindag2016}.
}

The introduction of prenatal sex determination fundamentally changed the 
relationship between son preference and birth spacing.
Couples, who before had shorter birth spacing because of son preference,
now have access to prenatal sex determination and sex-selective abortions,
and each abortion increases birth spacing by approximately a year.
The increase consists of three parts.
First, starting from the time of the abortion, the uterus needs at 
least two menstrual cycles to recover;  otherwise, the likelihood 
of spontaneous abortion increases substantially \citep{zhou00b}.
The second part is the waiting time to conception, which I set to six 
months \citep{Wang2003}.%
\footnote{
The waiting time to conception varies by woman, but even if it is, say, one 
month, the additional space between births would still be six months per abortion.
}
Finally, sex determination tests are reliable only from 3 months 
of gestation onwards.

@@

[To be added]

The reversal have some weird side effects.
Because there is now longer spacing after girls and this is when
we should observe sex selection we can get an less unequal 
sex ratio in the survey data than what we will eventually observe.

>>>>

Parents' spacing between births has long served as a measure of son preference 
\citep{Leung1988}.
Before prenatal sex determination became available, the only recourse for 
parents who wanted a son---but did not yet have one---was to have the next 
birth sooner.
Son preference is therefore often associated with shorter spacing after the 
births of girls than boys 
\citep{Das1987,Rahman1993,Pong1994,Haughton1996,Arnold1997,Soest2012,Rossi2015}.
Shorter spacing is, in turn, associated with worse health outcomes for girls 
and mothers 
\citep{arnold98,Conde-Agudelo2000,Whitworth2002,Razzaque2005,Rutstein2005,Conde-Agudelo2006}.%
\footnote{
Parents are also more likely to cease childbearing after the birth of 
a son than after a daughter 
\citep{ben-porath76b,Das1987,Arnold1997,clark00,filmer09,Basu2010,Altindag2016}.
}

But, the introduction of prenatal sex determination fundamentally changed the 
relationship between son preference and birth spacing.
Couples, who before had shorter birth spacing because of son preference,
now have access to prenatal sex determination and sex-selective abortions,
and each abortion increases birth spacing by approximately a year.
The increase consists of three parts.
First, starting from the time of the abortion, the uterus needs at 
least two menstrual cycles to recover;  otherwise, the likelihood 
of spontaneous abortion increases substantially \citep{zhou00b}.
The second part is the waiting time to conception, which I set to six 
months \citep{Wang2003}.%
\footnote{
The waiting time to conception varies by woman, but even if it is, say, one 
month, the additional space between births would still be six months per abortion.
}
Finally, sex determination tests are reliable only from 3 months 
of gestation onwards.

@@

The main question addressed here is the extent to which spacing patterns 
significantly changed as prenatal sex determination became available,
specifically, did the spacing after girls increase relative to the spacing
after boys?
The underlying idea is that in the absence of sex selection son 
preference leads to shorter spacing after the birth of a girl than after the 
birth of a boy, whereas son preference increases the spacing after the birth 
of a girl relative to after the birth of a boy when prenatal sex selection is 
available.
The four rounds of India's NFHS provide strong support for the hypothesis.

Censoring is an important issue when trying to understand the development
in sex ratio, especially when it comes to establishing whether the
use of sex selection is increasing or decreasing.
